\fontsize{16pt}{10pt}\selectfont\textbf{Structural controlability of pollination networks}

### 2015-12-05 paper/manuscript.Rnw
update manuscript and remove tex files

\input{preamble} % See preamble.tex to edit the overall layout

\begin{document}
\SweaveOpts{concordance=TRUE,prefix.string=figures/fig-}
\setkeys{Gin}{width=1\textwidth}

% title
\input{title_authors}
\maketitle

@@

There is no significant difference between the number of species needed to control the system in an invaded and an uninvaded ecosystem.

@@

ggplot(nr_sum) +
	geom_boxplot(aes(x = type, y = prop, fill = robustness)) +
	geom_vline(xintercept = seq(1.5, 5.5), linetype = 2) +
	xlab("link structure scheme") +
	ylab("species proportion") +
	my_theme + theme(legend.position = "right")
@
\caption{\label{fig:s_crit_prop} Proportion of species that are ordinary (when removed the number of species needed to control the network doesn't change), redundant (when removed less species are needed) and critical (when removed the number of species needed to control the network increasses). Schemes same as in \autoref{fig:n_species}}
\end{figure}

@@

ggplot(lr_sum) +
	geom_boxplot(aes(x = type, y = prop, fill = robustness)) +
	geom_vline(xintercept = seq(1.5, 5.5), linetype = 2) +
	xlab("link structure scheme") +
	ylab("interactions proportion") +
	my_theme + theme(legend.position = "right")
@
\caption{\label{fig:i_crit_prop} Proportion of interactions that are ordinary (when removed the number of species needed to control is reduced), redundant (when removed the number of species needed to control doesn't change) and critical (when removed the number of species needed to control the network increasses). Schemes same as in \autoref{fig:n_species}}
\end{figure}

@@

\citet{Ruths2014} proposed another clever way to characterise the control profile of a network by charactering the number of nodes that need to be controled among the ``mechanism'' that originates it.

@@

\bibliography{references}
\end{document}

>>>>

\documentclass[a4paper,10pt]{article}
% abstract + title + authors
\usepackage{abstract}
\renewcommand{\abstractnamefont}{\normalfont\bfseries}
\renewcommand{\abstracttextfont}{\normalfont\small\itshape} 
\usepackage{titlesec}
\usepackage{authblk}
\usepackage{datetime}
\newdateformat{usvardate}{
\monthname[\THEMONTH] \ordinal{DAY}, \THEYEAR}

% keywords
\providecommand{\keywords}[1]{\textbf{\textit{Keywords---}} #1}

% math
\usepackage{amssymb}
\usepackage{newtxmath}

% manuscript looking document
\usepackage{setspace}
\usepackage{lineno,xcolor}
\setlength{\parskip}{0.5em}

% comments

% graphics
\usepackage{graphicx}
\usepackage{float}

% references & links
\usepackage[
backend=biber,
style= authoryear,
citestyle = authoryear,
url=false,
doi=true,
isbn=false
]{biblatex}
\addbibresource{references.bib}

\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    citecolor=magenta,
    linkcolor=magenta
}
\usepackage{doi}
\begin{document}
\SweaveOpts{concordance=TRUE,prefix.string=figures/fig-}
\setkeys{Gin}{width=1\textwidth}

% title
% Article title
  \title{\fontsize{16pt}{10pt}\selectfont\textbf{Structural controlability of pollination networks}}

% Authors and Affiliations
  \author[1]{\large E. Fernando Cagua}
  \author[1]{\large Daniel B. Stouffer}
  \affil[1]{\normalsize Center of Integrative Ecology, School of Biological Sciences, University of Canterbury}

 \date{\small\usvardate\today}

\maketitle

@@

There was no significant difference between the number of species needed to control the system in an invaded and an uninvaded ecosystem.

When comparing the number of control nodes of the observed networks to their random counterparts an interesting pattern emegerges (\autoref{fig:random_n_species}). The number of control nodes is significantly larger than randomisations that mantain the degree distribution of the plants. However there is no significant difference when the randomisations mantain the degree of pollinators (or both plants and pollinators). This pattern is consistent across all schemes of link directionality. This suggest that the patterns of pollinator visitations are the structural characteristic that determines the controlability of the network. We probably should analyse other pollination networks to see if this pattern hold. We also should think about the implications of this finding.

\begin{figure}
\centering
<<fig = T, echo = F, width = 6, height = 3.5>>=
random_n <- read.csv("../data/random_n_driver.csv")
n_control <- n_driver %>%
	dplyr::select(-keep) %>%
	tidyr::separate(type, c("type", "keep"), "\\.", remove = T)

random_n_z <- random_n %>% 
	dplyr::group_by(net, type, keep, method) %>%
	dplyr::summarise(n_driver_u = mean(n_driver, na.rm = T),
									 n_driver_sd = sd(n_driver, na.rm = T)) %>%
	dplyr::full_join(n_control) %>% 
	dplyr::group_by() %>%
	dplyr::mutate(n_driver_z_score = (n_driver - n_driver_u)/n_driver_sd)

random_n_z %>% 
	ggplot(aes(x  = as.character(interaction(type, keep)), y = n_driver_z_score)) +
	geom_rect(ymin = -2, ymax = 2, xmin = -1, xmax = 10, fill = "grey90") +
	geom_hline(yintercept = 0, size = 0.4) +
	geom_vline(xintercept = seq(1.5, 5.5), linetype = 2) +
	geom_boxplot(aes(fill = method)) +
	scale_fill_discrete(name = "randomisations\nmantain\ndegree of", labels= c("plants", "pollinators", "both")) + 
	my_theme + 
	xlab("link structure scheme") +
	ylab("standard score") +
	theme(legend.position = "right",
				axis.line.x = element_blank(), axis.ticks.x = element_blank())
@
\caption{
\label{fig:random_n_species} 
Difference between the number of control nodes in the analysed networks versus randomisations of the same networks. Unshaded areas suggest a significant difference. Schemes same as in \autoref{fig:n_species}}
\end{figure}

It occured to me that it is possible that the patterns (and perhaps others observed like more phylogenetic signal in polinators than in plants) are a consequence of a greater diversity of pollinators in the interaction networks. More pollinators relative to plants imply that randomisations that mantain the degree distribution of pollinators are more constrained than those that only mantain the degree distribution of plants. It also means that statistical tests on that group have a more observations and therefore rendering significant results for smaller effect sizes. I cheked that with the 12 pollination networks of the study and there seems to be some evidence for that (\autoref{fig:zscore_ratio}). Perhaps it would be good to check with other networks to see if it holds.

\begin{figure}[h]
\centering
<<fig = T, echo = F, width = 6, height = 3.5>>=
net_info <- read.csv("../data/ntw_info.csv")

net_info %<>%
	dplyr::mutate(ratio_pol_pla = n_pol/n_pla)

random_n_z %>%
	dplyr::full_join(net_info) %>%
	dplyr::mutate(t = as.character(interaction(type, keep))) %>% 
	dplyr::filter(method == "c0", t != "BA.B") %>%
	ggplot(aes(x = ratio_pol_pla, y = n_driver_z_score)) +
	geom_point() +
	geom_smooth(method = "lm", alpha = 0.1, colour = "grey10") +
	xlab("ratio of number of pollinators to plants") +
	ylab("standard score") +
	my_theme + 
	facet_wrap(~t, scales = "free") + 
	theme(strip.background = element_blank())
@
\caption{
\label{fig:zscore_ratio} 
When considering only randomisations that mantain the degree distribution of plants, there seems to be a possitive relationship between the ratio of pollinators to plants and the Z-score of randomisations. The pattern is, however, not significant for the weighted schemes. The z-scores are identical for the AB.A and the BA.B schemes. Schemes same as in \autoref{fig:n_species}}
\end{figure}

@@

ggplot(nr_sum) +
	geom_boxplot(aes(x = type, y = prop, fill = robustness)) +
	geom_vline(xintercept = seq(1.5, 5.5), linetype = 2) +
	xlab("link structure scheme") +
	ylab("species proportion") +
	scale_fill_discrete(name = "nodes") +
	my_theme + 
	theme(legend.position = "right", 
				axis.line.x = element_blank(), 
				axis.ticks.x = element_blank())
@
\caption{\label{fig:s_crit_prop} Proportion of species that are ordinary (when removed the number of species needed to control the network doesn't change), redundant (when removed less species are needed) and critical (when removed the number of species needed to control the network increasses). Schemes same as in \autoref{fig:n_species}}
\end{figure}

@@

ggplot(lr_sum) +
	geom_boxplot(aes(x = type, y = prop, fill = robustness)) +
	geom_vline(xintercept = seq(1.5, 5.5), linetype = 2) +
	scale_fill_discrete(name = "links") +
	xlab("link structure scheme") +
	ylab("interactions proportion") +
	my_theme + theme(legend.position = "right")
@
\caption{\label{fig:i_crit_prop} Proportion of interactions that are ordinary (when removed the number of species needed to control is reduced), redundant (when removed the number of species needed to control doesn't change) and critical (when removed the number of species needed to control the network increasses). Schemes same as in \autoref{fig:n_species}}
\end{figure}

@@

\textcite{Ruths2014} proposed another clever way to characterise the control profile of a network by charactering the number of nodes that need to be controled among the ``mechanism'' that originates it. Similar as the maximum matching algorith, the idea comes from the concept that in the absense of cycles one node can at most control another one. Then the origin of control nodes can be splitted between nodes that need to be controled because they are a ``source node'' those that have only out links, control nodes associated to ``external dilations'' and those associated to ``internal dilations''.

It could be interesting to compare different networks using this framework. They argue that source dominated networks should be relatively easy to control in a top-down fashion. On the other hand internal dilation dominated networks represent the signature of a network that self-regulates and is not meant to be controled in that it's not designed to visit every possible state.

However as we can see here this framework is pretty useless when the direction of control is not perfectly known (as in most real wolrd networks really...). For instance when applying it to the data-set (\autoref{fig:dilation}), we see that when the control flows from pollinators to plants (AB.A) the network is source dominated simply because the pollinator diversity is greater than the plant's. When we asume that each guild can control each other (z-bi.all) the networks are internal dilation dominated because it's purely composed of cycles. However the weighted version might offer a meaningful possibility to develop the analysis.

@@

% \bibliography{references}
\printbibliography
\end{document}
